---
title: "Stats-Final"
author: "Syed Muhammad Hasan Zaidi"
date: "2023-12-02"
output: html_document
---
#I performed initial data cleaning using excel
```{r}
install.packages("data.table")
library(data.table)

data_2016 <- fread("C:\\Users\\hasan\\Downloads\\2016_brooklyn.csv")
data_2017 <- fread("C:\\Users\\hasan\\Downloads\\2017_brooklyn.csv")
data_2018 <- fread("C:\\Users\\hasan\\Downloads\\2018_brooklyn.csv")
data_2019 <- fread("C:\\Users\\hasan\\Downloads\\2019_brooklyn.csv")
data_2020 <- fread("C:\\Users\\hasan\\Downloads\\2020_brooklyn.csv")
```

```{r}
names(data_2016) <- c('borough','neighborhood','bldclasscat','taxclasscurr','block','lot','easement','bldclasscurr','address','aptnum','zip','resunits','comunits','totunits','landsqft','grosssqft','yrbuilt','taxclasssale','bldclasssale','price','date')

names(data_2017) <- c('borough','neighborhood','bldclasscat','taxclasscurr','block','lot','easement','bldclasscurr','address','aptnum','zip','resunits','comunits','totunits','landsqft','grosssqft','yrbuilt','taxclasssale','bldclasssale','price','date')

names(data_2018) <- c('borough','neighborhood','bldclasscat','taxclasscurr','block','lot','easement','bldclasscurr','address','aptnum','zip','resunits','comunits','totunits','landsqft','grosssqft','yrbuilt','taxclasssale','bldclasssale','price','date')

names(data_2019) <- c('borough','neighborhood','bldclasscat','taxclasscurr','block','lot','easement','bldclasscurr','address','aptnum','zip','resunits','comunits','totunits','landsqft','grosssqft','yrbuilt','taxclasssale','bldclasssale','price','date')

names(data_2020) <- c('borough','neighborhood','bldclasscat','taxclasscurr','block','lot','easement','bldclasscurr','address','aptnum','zip','resunits','comunits','totunits','landsqft','grosssqft','yrbuilt','taxclasssale','bldclasssale','price','date')


```

```{r}
combined_data <- rbind(data_2016, data_2017, data_2018, data_2019, data_2020)
total_rows <- nrow(combined_data)
print(total_rows)

na_count <- sum(is.na(combined_data$price))
na_count
```
#confirming data types
```{r}
column_types <- sapply(cleaned_data2, class)
result_table <- data.frame(Column_Names = names(column_types), Data_Types = column_types)
print(result_table)
```
#triming down the data set of any whitespaces 
```{r}
whitespace_check <- sapply(cleaned_data2, function(x) any(grepl("^\\s+|\\s+$", x)))
print(whitespace_check)
```

```{r}
hist(cleaned_data2$price, main = "Distribution of Price", xlab = "Price")
summary(cleaned_data2$price)

```
```{r}
cleaned_data2$price_bin <- cut(cleaned_data2$price, breaks = seq(0, max(cleaned_data2$price) + 10000, by = 10000), labels = FALSE)
cleaned_data2$price_bin <- cut(cleaned_data2$price, breaks = seq(10000, max(cleaned_data2$price) + 10000, by = 10000), labels = FALSE)


```

```{r}
cleaned_data2$price_bin <- cut(
  cleaned_data2$price,
  breaks = seq(0, ceiling(max(cleaned_data2$price) / 10000) * 10000, by = 10000),
  labels = seq(10000, ceiling(max(cleaned_data2$price) / 10000) * 10000, by = 10000),
  include.lowest = TRUE,
  right = FALSE
)
summary(cleaned_data2$price_bin)

```

```{r}
library(ggplot2)

# Count observations per price_bin
price_bin_counts <- table(cleaned_data2$price_bin)

# View the counts
print(price_bin_counts)

# Visualize the distribution
ggplot(cleaned_data2, aes(x = price_bin)) +
  geom_bar() +
  labs(x = "Price Bin", y = "Count") +
  ggtitle("Distribution of Price Bins")
```
```{r}
max_bin <- names(which.max(price_bin_counts))

# Display the bin with the highest count
print(paste("Bin with the highest count:", max_bin))
```
```{r}
Q1 <- quantile(cleaned_data$price, 0.25)
Q3 <- quantile(cleaned_data$price, 0.75)
IQR_val <- Q3 - Q1
upper_bound <- Q3 + 1.5 * IQR_val

# Remove upper bound outliers from 'price' column
cleaned_data3 <- cleaned_data2[cleaned_data2$price <= upper_bound, ]
```




```{r}
numeric_data <- cleaned_data3[sapply(cleaned_data3, is.numeric)]

```



```{r}
encoded_price_bin <- model.matrix(~ price_bin - 1, data = cleaned_data3)

```

```{r}
pca_result <- prcomp(cleaned_data3, center = TRUE, scale. = TRUE)

# Access the principal components
pcs <- pca_result$x

# Summary of PCA
summary(pca_result)
```
```{r}
na_counts <- colSums(is.na(cleaned_data2))

# Create a table displaying the count of NA values per column
na_table <- data.frame(Column_Names = names(na_counts), NA_Count = na_counts)
```
```{r}
na_table
```
```{r}
cleaned_data2$brook_neigh <- ifelse(is.na(cleaned_data2$brook_neigh), 'other', cleaned_data2$brook_neigh)

```

```{r}
cleaned_data2$price_bin <- as.numeric(cleaned_data2$price_bin)

```

```{r}
cleaned_data2$block_bin <- substr(cleaned_data2$block, 1, 2)

```

```{r}
cleaned_data2$zip_bin <- substr(cleaned_data2$zip, nchar(cleaned_data2$zip) - 1, nchar(cleaned_data2$zip))

```

```{r}
summary(cleaned_data2$lot)
min(cleaned_data2$lot, na.rm = TRUE)  
max(cleaned_data2$lot, na.rm = TRUE)  
mean(cleaned_data2$lot, na.rm = TRUE)
class(cleaned_data2$lot)
cleaned_data2$lot <- as.numeric(cleaned_data2$lot)
```
```{r}
cleaned_data2$lot_bin <- ceiling(cleaned_data2$lot / 10) * 10
class(cleaned_data2$lot_bin)

```
```{r}
min(cleaned_data2$grosssqft, na.rm = TRUE)  
max(cleaned_data2$grosssqft, na.rm = TRUE)  
mean(cleaned_data2$grosssqft, na.rm = TRUE)
class(cleaned_data2$grosssqft)
```
```{r}
cleaned_data2$gross_bin <- ceiling(cleaned_data2$grosssqft / 100) * 100

```


#sorting all dates to be in the same format
```{r}
library(lubridate)
combined_data$date <- parse_date_time(
  combined_data$date,
  orders = c("mdy", "dmy", "ymd"),
  quiet = TRUE
)

combined_data$date <- format(combined_data$date, "%m/%d/%y")

```

#adding columns. 1. Date
```{r}
combined_data$date <- mdy(combined_data$date)

combined_data$year <- year(combined_data$date)

get_quarter <- function(month) {
  ifelse(month %in% 1:3, "Qtr1",
         ifelse(month %in% 4:6, "Qtr2",
                ifelse(month %in% 7:9, "Qtr3", "Qtr4")))
}

combined_data$quarter <- get_quarter(month(combined_data$date))

```

#removing commas 
```{r}
library(dplyr)
library(stringr)

cleaned_data <- combined_data %>%
  mutate_all(~str_replace_all(., ",", ""))

```

#Changing Data Types```
```{r}
na_count <- sum(is.na(cleaned_data$price))
na_count
```
```{r}
cleaned_data$price <- gsub("\\$", "", cleaned_data$price)

```

```{r}
na_count <- sum(is.na(cleaned_data$price))
na_count
```

```{r}
cleaned_data$price <- as.integer(cleaned_data$price)
cleaned_data$grosssqft <- as.numeric(cleaned_data$grosssqft)
cleaned_data$landsqft <- as.numeric(cleaned_data$landsqft)

```

```{r}
cleaned_data2 <- cleaned_data
```

```{r}
cleaned_data2 <- cleaned_data2[!is.na(cleaned_data2$grosssqft), ]
```


```{r}
neighborhood_counts <- table(cleaned_data2$neighborhood)
print(neighborhood_counts)

```
```{r}
year_counts <- table(cleaned_data2$year)
print(year_counts)

```
```{r}
zip_counts <- table(cleaned_data2$zip)
print(zip_counts)

```

```{r}
cleaned_data2$sqrt_price <- sqrt(cleaned_data2$price)
cleaned_data2$sqrt_grosssqft <- sqrt(cleaned_data2$grosssqft)
cleaned_data2$sqrt_landsqft <- sqrt(cleaned_data2$landsqft)
```

```{r}
cleaned_data2$log_price <- log(cleaned_data2$price)
cleaned_data2$log_grosssqft <- log(cleaned_data2$grosssqft)
cleaned_data2$log_landsqft <- log(cleaned_data2$landsqft)
```

```{r}
library(dplyr)

# Your matched neighborhood data
matches <- tibble::tribble(
  ~List2Neighborhood, ~List1Heading,
  "BATH BEACH", "Southern Brooklyn",
  "BAY RIDGE", "Southwestern Brooklyn",
  "BEDFORD STUYVESANT", "Northern Brooklyn",
  # Add other matched neighborhoods here...
)

# Creating the new column 'brook_neigh'
cleaned_data2 <- cleaned_data2 %>%
  left_join(matches, by = c("neighborhood" = "List2Neighborhood")) %>%
  mutate(brook_neigh = case_when(
    !is.na(List1Heading) ~ List1Heading,
    TRUE ~ NA_character_  # If no match is found
  ))
```

```{r}
hist(cleaned_data2$price, main = "Distribution of Price", xlab = "Price")

```

```{r}
min(cleaned_data2$gross_bin, na.rm = TRUE)  
max(cleaned_data2$gross_bin, na.rm = TRUE)  
mean(cleaned_data2$gross_bin, na.rm = TRUE)
```
```{r}
mean_gross <- mean(cleaned_data2$gross_bin)
sd_gross <- sd(cleaned_data2$gross_bin)
upper_bound <- mean_gross + 3 * sd_gross
lower_bound <- mean_gross - 3 * sd_gross
outliers <- cleaned_data2$gross_bin[cleaned_data2$gross_bin > upper_bound | cleaned_data2$gross_bin < lower_bound]
num_outliers <- length(outliers)
cat("Number of outliers:", num_outliers, "\n")

```
```{r}

cleaned_data2 <- subset(cleaned_data2, !(gross_bin > upper_bound | gross_bin < lower_bound))
```

```{r}
class(cleaned_data2$zip_bin)

```
#Final Data
```{r}
num_rows <- nrow(cleaned_data2)
print(num_rows)
```

#Model
```{r}
lm_model_price_bin <- lm(price ~ factor(brook_neigh)   + factor(zip_bin) + gross_bin , data = cleaned_data2)
```


```{r}
summary(lm_model_price_bin)

# RMSE calculation
predicted <- predict(lm_model_price_bin)
actual <- cleaned_data2$price
rmse <- sqrt(mean((actual - predicted)^2))
print(rmse)
```
# R^2 = 0.5585
# DF 41
#RSME 463979.9

```{r}
saveRDS(list(model=lm_model_price_bin, data=cleaned_data2), file='HasanFinal.RDS') 
```

```{r}
saveRDS(lm_model_price_bin, file = "Hasan_Final.rds")

```

```{r}
getwd()
```
```{r}
list.files()

list.files(pattern = "lm_model_price_bin")

```

```{r}
setwd("~/Desktop")

```

```{r}
list.files(pattern = "lm_model_price_bin")

```

```{r}
print(getwd())

```
```{r}

```



